---
title: "Home Video Feasibility"
author: "Megan"
date: "2025-04-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(dplyr)
library(tidyverse)
library(ggplot2)
```
# Resource for selecting test 
https://stats.oarc.ucla.edu/other/mult-pkg/whatstat/ 

# set output directory 
```{r}
# analysis folder 
analysis_version <- '007'
```

```{r}
output_dir <- file.path("C:/Users/mmccu/Box/MM_Personal/5_Projects/BoveLab/3_Data_and_Code/gait_bw_zeno_home_analysis",
                    analysis_version, 
                    "004c_home_video_feasibility")

ifelse(!dir.exists(output_dir), 
       dir.create(output_dir), 
       "directory already exists") 
```

# Load data 
files are first visit files from 004 demographics jupyternotebook files 
```{r}
# Participants from all BW data - clean (has all data in range)
clean_bw_ms_path <- file.path("C:/Users/mmccu/Box/MM_Personal/5_Projects/BoveLab/3_Data_and_Code/gait_bw_zeno_home_analysis",
                              analysis_version, 
                              "004_demographics_feasibility", 
                              "clean_bw_ms_all_first_visit.csv")
print(clean_bw_ms_path)
clean_bw_ms_df <- read.csv(clean_bw_ms_path)
```

```{r}
head(clean_bw_ms_df)
```

```{r}
# Participants consented to home videos - clean (has all data in range)
clean_hv_consent_path <- file.path("C:/Users/mmccu/Box/MM_Personal/5_Projects/BoveLab/3_Data_and_Code/gait_bw_zeno_home_analysis",
                                   analysis_version, 
                                   "004_demographics_feasibility", 
                                   "home_video_consent_clean_first_visit.csv")
print(clean_hv_consent_path)
clean_hv_consent_df <- read.csv(clean_hv_consent_path)
```
```{r}
head(clean_hv_consent_df)
```


```{r}
# Participants who sent home videos - clean (has all data in range)
clean_hv_video_path <- file.path("C:/Users/mmccu/Box/MM_Personal/5_Projects/BoveLab/3_Data_and_Code/gait_bw_zeno_home_analysis",
                                   analysis_version, 
                                   "004_demographics_feasibility", 
                                    "clean_home_first_visit.csv")

print(clean_hv_video_path)
clean_hv_video_df <- read.csv(clean_hv_video_path)
```
```{r}
head(clean_hv_video_df)
```



# Format data 

 
```{r}
factor_vars <- function(df){
  df <- df %>% 
    mutate_at(c("clean_Sex", "bingoEHR_DX_MS.DX", 
              "demographic_diagnosis", "race_ethnicity_clean", 
              "ms_dx_condensed"), 
            as.factor)
  
  # set levels 
  df$race_ethnicity_clean <- factor(df$race_ethnicity_clean, 
                                    levels = c('White Non Hispanic', 
                                               'Asian', 
                                               'Black Or African American',
                                               'Hispanic or Latino',
                                               'Other/Unknown/Declined'), 
                                           ordered = FALSE)
  
  # merge ms_dx into RRMS or other 
  df <- df %>%
    mutate(ms_dx_condensed_2 = case_when(
      ms_dx_condensed == "RRMS" ~ "RRMS",
      ms_dx_condensed == "Progressive MS" ~ "Progressive or Not Specified",
      ms_dx_condensed == "MS, Subtype Not Specified" ~ "Progressive or Not Specified"
      ))
  
  df$ms_dx_condensed_2 <- factor(df$ms_dx_condensed_2, 
                               levels = c('RRMS',
                                          'Progressive or Not Specified'),
                               ordered = FALSE)
  
  
  # merge sex and factor 
  df <- df %>%
    mutate(clean_Sex_2 = case_when(
      clean_Sex == "Male" ~ "Male or Non-Binary",
      clean_Sex == "Non-Binary" ~ "Male or Non-Binary",
      clean_Sex == "Female" ~ "Female"
      ))
  
  df$clean_Sex_2 <- factor(df$clean_Sex_2, 
                          levels = c('Female',
                                     'Male or Non-Binary'),
                               ordered = FALSE)

  return(df)
}

clean_bw_ms_df <- factor_vars(clean_bw_ms_df)
clean_hv_consent_df <- factor_vars(clean_hv_consent_df)
clean_hv_video_df <- factor_vars(clean_hv_video_df)


```



# Create new data frame - not consent 
Need mutually exclusive groups for test 
find people who did not consent, who consented and did not sent videos 
```{r}
# bw participants who did not consent to home videos 
no_hv_consent_df <- clean_bw_ms_df %>% 
  filter(!(bw_id %in% clean_hv_consent_df$bw_id))


# check 
table(clean_bw_ms_df$clean_Sex_2)
table(no_hv_consent_df$clean_Sex_2)
table(clean_hv_consent_df$clean_Sex_2)

```
this might be mismatched until running final report - some sent videos recently 
```{r}
# participants who consented, but did not sent a video 
no_video_df <- clean_hv_consent_df %>% 
  filter(!(bw_id %in% clean_hv_video_df$bw_id))

# mutually exclusive
duplicates <- clean_hv_video_df$bw_id[clean_hv_video_df$bw_id %in% no_video_df$bw_id]
duplicates
```

# Categorical variables 
Higher X-squared = greater difference between groups 
p > .05 - not statistically significant 

## Functions 

```{r}
print_save_xsq_results <- function(xsq_results, column){ 
  
  print(xsq_results)
  
  print('------ expected ---------') 
  print(xsq_results$expected)
  
  print('------ residuals ---------') 
  print(xsq_results$residuals)
  
  # save row for data frame 
  results_row <- data.frame(column = column, 
                            statistic = xsq_results$statistic,
                            p_value = xsq_results$p.value, 
                            degrees_freedom = xsq_results$parameter, 
                            method = xsq_results$method)
  return(results_row)
  
  
  }
```


```{r}
print_save_fisher_results <- function(fisher_results, column){

  results_row <- data.frame(column = column, 
                            statistic = NaN,
                            p_value = fisher_results$p.value, 
                            degrees_freedom = NaN,
                            method = fisher_results$method)
  
  return(results_row)
  
  }
```


## Sex
```{r}
# consented to home videos vs did not consent to home videos 

# any missing? 
print('----- missing ------')
sum(is.na(no_hv_consent_df$clean_Sex_2))
sum(is.na(clean_hv_consent_df$clean_Sex_2))

# format table 
consent_sex_tab <- rbind(table(no_hv_consent_df$clean_Sex_2), table(clean_hv_consent_df$clean_Sex_2))
dimnames(consent_sex_tab)[1] <- list("Participants:" = c("Did not consent", "Consented"))
consent_sex_tab

# chi square test 
consent_sex_xsq <- chisq.test(consent_sex_tab, correct = FALSE)
consent_sex_xsq_row <- print_save_xsq_results(consent_sex_xsq, column = 'Sex')
```
```{r}
# any missing? 
print('----- missing ------')
sum(is.na(no_video_df$clean_Sex_2))
sum(is.na(clean_hv_video_df$clean_Sex_2))

# sent home video vs did not send home video
video_sex_tab <- rbind(table(no_video_df$clean_Sex_2), table(clean_hv_video_df$clean_Sex_2))


dimnames(video_sex_tab)[1] <- list("Participants:" = c("Did not sent video", "Sent video"))
video_sex_tab

# chi square test 
video_sex_xsq <- chisq.test(video_sex_tab, correct = FALSE)
video_sex_xsq_row <- print_save_xsq_results(video_sex_xsq, column = 'Sex')
```
## MS Subtype 
```{r}
# consented to home videos vs did not consent to home videos 

print('----- missing ------')
sum(is.na(no_hv_consent_df$ms_dx_condensed_2))
sum(is.na(clean_hv_consent_df$ms_dx_condensed_2))

# format table 
consent_dx_tab <- rbind(table(no_hv_consent_df$ms_dx_condensed_2), table(clean_hv_consent_df$ms_dx_condensed_2))
dimnames(consent_dx_tab)[1] <- list("Participants:" = c("Did not consent", "Consented"))

consent_dx_tab

# chi square test 
consent_dx_xsq <- chisq.test(consent_dx_tab, correct = FALSE)
consent_dx_xsq_row <- print_save_xsq_results(consent_dx_xsq, column = 'MS DX')

```
```{r}
# sent home video vs did not send home video

print('----- missing ------')
sum(is.na(no_video_df$ms_dx_condensed_2))
sum(is.na(clean_hv_video_df$ms_dx_condensed_2))

video_dx_tab <- rbind(table(no_video_df$ms_dx_condensed_2), table(clean_hv_video_df$ms_dx_condensed_2))


dimnames(video_dx_tab)[1] <- list("Participants:" = c("Did not sent video", "Sent video"))
video_dx_tab

# chi square test 
video_dx_fisher <- fisher.test(video_dx_tab)
video_dx_fisher
video_dx_fisher_row <- print_save_fisher_results(video_dx_fisher, 'MS DX')
```
## Race and Ethnicity 
Question - should I group into white non hispanic / other for 2x2?
```{r}
# consented to home videos vs did not consent to home videos 

print('----- missing ------')
sum(is.na(no_hv_consent_df$race_ethnicity_clean))
sum(is.na(clean_hv_consent_df$race_ethnicity_clean))

# format table 
consent_race_tab <- rbind(table(no_hv_consent_df$race_ethnicity_clean), table(clean_hv_consent_df$race_ethnicity_clean))
dimnames(consent_race_tab)[1] <- list("Participants:" = c("Did not consent", "Consented"))

consent_race_tab

# Fishers test 
consent_race_fisher <- fisher.test(consent_race_tab)
consent_race_fisher
consent_race_fisher_row <- print_save_fisher_results(consent_race_fisher, 'Race and Ethnicity')
```


```{r}
# sent home video vs did not send home video
print('----- missing ------')
sum(is.na(no_video_df$race_ethnicity_clean))
sum(is.na(clean_hv_video_df$race_ethnicity_clean))


video_race_tab <- rbind(table(no_video_df$race_ethnicity_clean), table(clean_hv_video_df$race_ethnicity_clean))


dimnames(video_race_tab)[1] <- list("Participants:" = c("Did not sent video", "Sent video"))
video_race_tab
video_race_fisher <- fisher.test(video_race_tab)
video_race_fisher
video_race_fisher_row <- print_save_fisher_results(video_race_fisher, 'Race and Ethnicity')

```

## Bind all rows together and save 
consent_sex_xsq_row with all other rows

```{r}
# consent vs no consent to home videos 
consent_cat_results <- rbind(consent_sex_xsq_row, consent_dx_xsq_row, consent_race_fisher_row)
consent_cat_results$p_value <- round(consent_cat_results$p_value, 3)
consent_cat_results
write.csv(consent_cat_results, file.path(output_dir, 'results_consent_vs_not_categorical.csv'))
```

```{r}
# sent home videos vs did not sent home videos 
video_cat_results <- rbind(video_sex_xsq_row, video_dx_fisher_row, video_race_fisher_row)
video_cat_results$p_value <- round(video_cat_results$p_value, 3)
video_cat_results
write.csv(video_cat_results, file.path(output_dir, 'results_video_vs_no_video_categorical.csv'))
```


## Categorical summary 
4/28/2025 - There was no significant difference in sex, MS diagnosis, or race between indiividuals that di and did not consent to home videos. 

Within individuals who consented to home videos, there was no signficiant difference in sex, MS diagnosis, or race between individuals that sent or did not send videos back to the study team. 

# Continuous Variables
Unpaired t-test: comparing two groups

## T-test assumptions 
1. Simple random sample 

2. Data approximately normally distributed in both groups: test = histogram and shapiro wilk test. Normal if p > 0.05. If not normal, Mann-Whitney/ Wilcoxon Rank Sum test for unpaired data  (or ordinal I think)

3. Variance of two groups is similar: test = variance ratio test.Consider equal if p > 0.05 If unequal variance, Weltche's t-test (t.test with var.equal argument = FALSE). 

## Functions 
```{r}
# summarize continuous ordinal data 
summarise_continous <- function(data, column) {
  
  # save row for data frame 
  summary_row <- data.frame(column = column, 
                            mean = round(mean(data, na.rm = TRUE), 2),
                            median = round(median(data, na.rm = TRUE), 2), 
                            q1 = round(quantile(data, probs = 0.25, na.rm = TRUE), 2),
                            q3 = round(quantile(data, probs = 0.75, na.rm = TRUE), 2),
                            sd = round(sd(data, na.rm = TRUE), 2), 
                            not_missing = sum(!is.na(data)),
                            missing = sum(is.na(data)))
  
  
  return(summary_row)
  
}
```

```{r}
save_test_results <- function(ttest_wilcox_results, column){ 
  
  results_row <- data.frame(column_name = column, 
                            statistic = ttest_wilcox_results$statistic,
                            p_value = ttest_wilcox_results$p.value, 
                            conf_int_1 = ttest_wilcox_results$conf.int[1],
                            conf_int_2 = ttest_wilcox_results$conf.int[2],
                            alternative = ttest_wilcox_results$alternative, 
                            method = ttest_wilcox_results$method, 
                            data_name = ttest_wilcox_results$data.name)
  
  return(results_row)
  
  }
```

## Age 
```{r}
# consent vs did not consent 
# summary 
sum_age_no_consent <- summarise_continous(no_hv_consent_df$clean_Age, 'Age')
sum_age_no_consent

sum_age_consent <- summarise_continous(clean_hv_consent_df$clean_Age, 'Age')
sum_age_consent

# compare two variances - welch's test if p < 0.05 
var.test(no_hv_consent_df$clean_Age, clean_hv_consent_df$clean_Age)

# histograms  
par(mfrow=c(1,2))
hist(no_hv_consent_df$clean_Age, probability =  FALSE)
hist(clean_hv_consent_df$clean_Age, probability =  FALSE)

# Shapiro-wilk tet of normality 
shapiro.test(no_hv_consent_df$clean_Age)
shapiro.test(clean_hv_consent_df$clean_Age)

# 4/28 - no consent group not normal, using Mann-Whitney/ Wilcoxon Rank Sum test 
consent_age_wilcox <- wilcox.test(x = no_hv_consent_df$clean_Age, 
                                  y = clean_hv_consent_df$clean_Age, 
                                  alternative = c('two.sided'), 
                                  paired = FALSE, 
                                  conf.int = TRUE) 

consent_age_results_row <- save_test_results(consent_age_wilcox, 'Age')
consent_age_results_row
```
```{r}
# sent or did not send video 

# summary 
sum_age_no_video <- summarise_continous(no_video_df$clean_Age, 'Age')
sum_age_no_video

sum_age_video <- summarise_continous(clean_hv_video_df$clean_Age, 'Age')
sum_age_video

# compare two variances - welch's test if p < 0.05 
var.test(no_video_df$clean_Age, clean_hv_video_df$clean_Age)

# histograms  
par(mfrow=c(1,2))
hist(no_video_df$clean_Age, probability =  FALSE)
hist(clean_hv_video_df$clean_Age, probability =  FALSE)

# Shapiro-wilk tet of normality 
shapiro.test(no_video_df$clean_Age)
shapiro.test(clean_hv_video_df$clean_Age)

# test - 4/28 - both shapiro tests > 0.05, t-test 
video_age_ttest <- t.test(x = no_video_df$clean_Age, 
                          y = clean_hv_video_df$clean_Age, 
                          var.equal = TRUE, 
                          alternative = c("two.sided"))

video_age_ttest 
video_age_results_row <- save_test_results(video_age_ttest, 'Age')

```

## Disease Duration 
```{r}
# consent vs did not consent 
# summary 
sum_dur_no_consent <- summarise_continous(no_hv_consent_df$demoEHR_DiseaseDuration,
                                          'Disease Duration')
sum_dur_no_consent

sum_dur_consent <- summarise_continous(clean_hv_consent_df$demoEHR_DiseaseDuration,
                                       'Disease Duration')
sum_dur_consent

# compare two variances - welch's test if p < 0.05 
var.test(no_hv_consent_df$demoEHR_DiseaseDuration, clean_hv_consent_df$demoEHR_DiseaseDuration)

# histograms  
par(mfrow=c(1,2))
hist(no_hv_consent_df$demoEHR_DiseaseDuration, probability =  FALSE)
hist(clean_hv_consent_df$demoEHR_DiseaseDuration, probability =  FALSE)

# Shapiro-wilk tet of normality 
shapiro.test(no_hv_consent_df$demoEHR_DiseaseDuration)
shapiro.test(clean_hv_consent_df$demoEHR_DiseaseDuration)

# test 
# 4/28 - neither normal, mann whitney 
consent_dur_wilcox <- wilcox.test(x = no_hv_consent_df$demoEHR_DiseaseDuration, 
                                  y = clean_hv_consent_df$demoEHR_DiseaseDuration, 
                                  alternative = c('two.sided'), 
                                  paired = FALSE, 
                                  conf.int = TRUE) 
consent_dur_wilcox
consent_dur_results_row <- save_test_results(consent_dur_wilcox, 'Disease Duration')
```

```{r}
# sent video vs did not send video 
# summary 
sum_dur_no_video <- summarise_continous(no_video_df$demoEHR_DiseaseDuration,
                                          'Disease Duration')
sum_dur_no_video

sum_dur_video <- summarise_continous(clean_hv_video_df$demoEHR_DiseaseDuration,
                                       'Disease Duration')
sum_dur_video

# compare two variances - welch's test if p < 0.05 
var.test(no_video_df$demoEHR_DiseaseDuration, 
         clean_hv_video_df$demoEHR_DiseaseDuration)

# histograms  
par(mfrow=c(1,2))
hist(no_video_df$demoEHR_DiseaseDuration, probability =  FALSE)
hist(clean_hv_video_df$demoEHR_DiseaseDuration, probability =  FALSE)

# Shapiro-wilk tet of normality 
shapiro.test(no_video_df$demoEHR_DiseaseDuration)
shapiro.test(clean_hv_video_df$demoEHR_DiseaseDuration)

# test 
# 4/28 - not normal, using wilcox test 
video_dur_wilcox <- wilcox.test(x = no_video_df$demoEHR_DiseaseDuration,
                                y = clean_hv_video_df$demoEHR_DiseaseDuration, 
                                alternative = c('two.sided'), 
                                paired = FALSE,
                                conf.int = TRUE) 
video_dur_wilcox
video_dur_results_row <- save_test_results(video_dur_wilcox, 'Disease Duration')
```
## T25FW
```{r}
# consent vs no consent 
# summary 
sum_t25fw_no_consent <- summarise_continous(no_hv_consent_df$clean_T25FW_Avg,
                                          'T25FW')
sum_t25fw_no_consent

sum_t25fw_consent <- summarise_continous(clean_hv_consent_df$clean_T25FW_Avg,
                                         'T25FW')
sum_t25fw_consent

# compare two variances - welch's test if p < 0.05 
var.test(no_hv_consent_df$clean_T25FW_Avg, 
         clean_hv_consent_df$clean_T25FW_Avg)

# histograms  
par(mfrow=c(1,2))
hist(no_hv_consent_df$clean_T25FW_Avg, probability =  FALSE)
hist(clean_hv_consent_df$clean_T25FW_Avg, probability =  FALSE)

# Shapiro-wilk tet of normality 
shapiro.test(no_hv_consent_df$clean_T25FW_Avg)
shapiro.test(clean_hv_consent_df$clean_T25FW_Avg)

# test 
# 4/28 - wilcox, not normal 
consent_t25fw_wilcox <- wilcox.test(x = no_hv_consent_df$clean_T25FW_Avg,
                                    y = clean_hv_consent_df$clean_T25FW_Avg,
                                    alternative = c('two.sided'), 
                                    paired = FALSE,
                                    conf.int = TRUE)

consent_t25fw_wilcox
consent_t25fw_results_row <- save_test_results(consent_t25fw_wilcox, 'T25FW')

```

```{r}
# home video vs no video 
# summary 
sum_t25fw_no_video <- summarise_continous(no_video_df$clean_T25FW_Avg,
                                          'T25FW')
sum_t25fw_no_video

sum_t25fw_video <- summarise_continous(clean_hv_video_df$clean_T25FW_Avg,
                                       'T25FW')
sum_t25fw_video

# compare two variances - welch's test if p < 0.05 
var.test(no_video_df$clean_T25FW_Avg, 
         clean_hv_video_df$clean_T25FW_Avg)

# histograms  
par(mfrow=c(1,2))
hist(no_video_df$clean_T25FW_Avg, probability =  FALSE)
hist(clean_hv_video_df$clean_T25FW_Avg, probability =  FALSE)

# Shapiro-wilk tet of normality 
shapiro.test(no_video_df$clean_T25FW_Avg)
shapiro.test(clean_hv_video_df$clean_T25FW_Avg)

# test 
# 4/28 - wilcox, not normal 
video_t25fw_wilcox <- wilcox.test(x = no_video_df$clean_T25FW_Avg,
                                  y = clean_hv_video_df$clean_T25FW_Avg,
                                  alternative = c('two.sided'), 
                                  paired = FALSE,
                                  conf.int = TRUE)
video_t25fw_wilcox
video_t25fw_results_row <- save_test_results(video_t25fw_wilcox, 'T25FW')
```


## Education 
```{r}
# consent vs no consent 
# summary 
sum_edu_no_consent <- summarise_continous(no_hv_consent_df$tc_Examinee_Education,
                                          'Education')
sum_edu_no_consent

sum_edu_consent <- summarise_continous(clean_hv_consent_df$tc_Examinee_Education,
                                         'Education')
sum_edu_consent

# compare two variances - welch's test if p < 0.05 
var.test(no_hv_consent_df$tc_Examinee_Education, clean_hv_consent_df$tc_Examinee_Education)

# histograms  
par(mfrow=c(1,2))
hist(no_hv_consent_df$tc_Examinee_Education, probability =  FALSE)
hist(clean_hv_consent_df$tc_Examinee_Education, probability =  FALSE)

# Shapiro-wilk tet of normality 
shapiro.test(no_hv_consent_df$tc_Examinee_Education)
shapiro.test(clean_hv_consent_df$tc_Examinee_Education)

# test 
# 4/28- wilcoxin not normal 
consent_edu_wilcox <- wilcox.test(x = no_hv_consent_df$tc_Examinee_Education,
                                  y = clean_hv_consent_df$tc_Examinee_Education,
                                  alternative = c('two.sided'), 
                                  paired = FALSE,
                                  conf.int = TRUE)
consent_edu_wilcox
consent_edu_results_row <- save_test_results(consent_edu_wilcox, 'Education')
```
```{r}
# home video vs not home video 
# summary 
sum_edu_no_video <- summarise_continous(no_video_df$tc_Examinee_Education,
                                        'Education')
sum_edu_no_video

sum_edu_video <- summarise_continous(clean_hv_video_df$tc_Examinee_Education,
                                     'Education')
sum_edu_video

# compare two variances - welch's test if p < 0.05 
var.test(no_video_df$tc_Examinee_Education, 
         clean_hv_video_df$tc_Examinee_Education)

# histograms  
par(mfrow=c(1,2))
hist(no_video_df$tc_Examinee_Education, probability =  FALSE)
hist(clean_hv_video_df$tc_Examinee_Education, probability =  FALSE)

#Shapiro-wilk tet of normality 
shapiro.test(no_video_df$tc_Examinee_Education)
shapiro.test(clean_hv_video_df$tc_Examinee_Education)

# test 
# 4/28 - wilcoxon not normal 
video_edu_wilcox <- wilcox.test(x = no_video_df$tc_Examinee_Education,
                                  y = clean_hv_video_df$tc_Examinee_Education,
                                  alternative = c('two.sided'), 
                                  paired = FALSE,
                                  conf.int = TRUE)
video_edu_wilcox
video_edu_results_row <- save_test_results(video_edu_wilcox, 'Education')
```

## EDSS 
Ordinal - use wilcoxin mann whitney 
```{r}
# consent vs no consent 
# summary 
sum_edss_no_consent <- summarise_continous(no_hv_consent_df$clean_EDSS,
                                          'EDSS')
sum_edss_no_consent

sum_edss_consent <- summarise_continous(clean_hv_consent_df$clean_EDSS,
                                         'EDSS')
sum_edss_consent

# histograms  
par(mfrow=c(1,2))
hist(no_hv_consent_df$clean_EDSS, probability =  FALSE)
hist(clean_hv_consent_df$clean_EDSS, probability =  FALSE)

# ordinal - 
consent_edss_wilcox <- wilcox.test(x = no_hv_consent_df$clean_EDSS,
                                  y = clean_hv_consent_df$clean_EDSS,
                                  alternative = c('two.sided'), 
                                  paired = FALSE,
                                  conf.int = TRUE)
consent_edss_wilcox
consent_edss_results_row <- save_test_results(consent_edss_wilcox, 'EDSS')
```
```{r}
# home video vs not home video 
# summary 
sum_edss_no_video <- summarise_continous(no_video_df$clean_EDSS,
                                        'EDSS')
sum_edss_no_video

sum_edss_video <- summarise_continous(clean_hv_video_df$clean_EDSS,
                                     'EDSS')
sum_edss_video

# histograms  
par(mfrow=c(1,2))
hist(no_video_df$clean_EDSS, probability =  FALSE)
hist(clean_hv_video_df$clean_EDSS, probability =  FALSE)


# test 
video_edss_wilcox <- wilcox.test(x = no_video_df$clean_EDSS,
                                  y = clean_hv_video_df$clean_EDSS,
                                  alternative = c('two.sided'), 
                                  paired = FALSE,
                                  conf.int = TRUE)
video_edss_wilcox
video_edss_results_row <- save_test_results(video_edss_wilcox, 'EDSS')

```


## Bind all results 
```{r}
#  consent 
# consent - no 
consent_no_df <- rbind(sum_age_no_consent, 
                      sum_dur_no_consent, 
                      sum_t25fw_no_consent,
                      sum_edu_no_consent,
                      sum_edss_no_consent) %>% 
  mutate('group_name' = 'Did not consent')
                       
# consent - yes 
consent_summary_yes <- rbind(sum_age_consent,
                             sum_dur_consent,
                             sum_t25fw_consent,
                             sum_edu_consent,
                             sum_edss_consent) %>% 
  mutate('group_name' = 'Consented')

# consent all 
consent_summary_all <- rbind(consent_no_df, consent_summary_yes) 
consent_summary_all <- consent_summary_all[order(consent_summary_all$column), ]
consent_summary_all
write.csv(consent_summary_all, file.path(output_dir, 'summary_consent_cont.csv'))
```


```{r}
# video or no video 
# video - no 
video_no_df <- rbind(sum_age_no_video,
                     sum_dur_no_video,
                     sum_t25fw_no_video,
                     sum_edu_no_video,
                     sum_edss_no_video) %>% 
  mutate('group_name' = 'Did not sent video')

# video - yes 
video_yes_df <- rbind(sum_age_video,
                      sum_dur_video,
                      sum_t25fw_video,
                      sum_edu_video,
                      sum_edss_video) %>% 
  mutate('group_name' = 'Sent video')

# video all 
video_summary_all <- rbind(video_no_df, video_yes_df)
video_summary_all <- video_summary_all[order(video_summary_all$column), ]
video_summary_all 
write.csv(video_summary_all, file.path(output_dir, 'summary_video_summary_cont.csv'))
```


```{r}
# test results 
# consent 
results_consent_cont <- rbind(consent_age_results_row,
                              consent_dur_results_row,
                              consent_t25fw_results_row,
                              consent_edu_results_row,
                              consent_edss_results_row) 
results_consent_cont
write.csv(results_consent_cont, file.path(output_dir, 'results_consent_vs_not_cont.csv'))

# video 

results_video_cont <- rbind(video_age_results_row, 
                            video_dur_results_row,
                            video_t25fw_results_row,
                            video_edu_results_row,
                            video_edss_results_row)
results_video_cont
write.csv(results_video_cont, file.path(output_dir, 'results_video_vs_no_video_cont.csv'))

```

