---
title: "Longitudal video metric analysis"
author: "Megan"
date: "2025-05-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggplot2)
library(psych)
library(broom.mixed)
```


# set output directory 
```{r}
# analysis folder 
analysis_version <- '011'
```


```{r}
long_output_dir <- file.path("C:/Users/mmccu/Box/MM_Personal/5_Projects/BoveLab/3_Data_and_Code/gait_bw_zeno_home_analysis",
                    analysis_version, 
                    "006_longitudinal_v3")

ifelse(!dir.exists(long_output_dir), 
       dir.create(long_output_dir), 
       "directory already exists") 


pca_output_dir <- file.path("C:/Users/mmccu/Box/MM_Personal/5_Projects/BoveLab/3_Data_and_Code/gait_bw_zeno_home_analysis",
                    analysis_version, 
                    "006_pca")

ifelse(!dir.exists(pca_output_dir), 
       dir.create(pca_output_dir), 
       "directory already exists") 
```

# Load FW video data from participants with baseline and 1 year visits 
Formatted and filtered in jupyter notebook
```{r}
zv_fw_base_yr1_path <- file.path("C:/Users/mmccu/Box/MM_Personal/5_Projects/BoveLab/3_Data_and_Code/gait_bw_zeno_home_analysis",
                       analysis_version, 
                       "006_longitudinal", 
                       "fw_base_yr2_1_all.csv")

zv_fw_base_yr1_df <- read.csv(zv_fw_base_yr1_path)
```


# Make long dataset with subset of cols for analysis  
Select variables for analysis  
```{r}

zv_fw_base_yr1_long <- zv_fw_base_yr1_df %>% 
  select(bw_id, redcap_event_name, video_id_date_name_pose_zv, id_date_pose_zv, 
         delta_pix_h_rel_median_pose_zv, stride_time_median_sec_pose_zv, 
         mean_cadence_step_per_min_pose_zv, stride_width_median_cm_pose_zv,
         trialdate, clean_T25FW_Avg, clean_EDSS, msfcEHR_T25FW.SPEED.TRIAL.1.vDate.Diff, 
         demoEHR_EDSS_dateDiff) %>% 
  mutate(redcap_event_clean = ifelse(redcap_event_name == 'Brainwalk: Baseline visit (Arm 1: Baseline visit)',
                                   'baseline',
                                   'year2')) 

```

## Check date diffs clinical outcomes 

T25FW
```{r}
p1 <- ggplot(data = zv_fw_base_yr1_long, aes(x = msfcEHR_T25FW.SPEED.TRIAL.1.vDate.Diff)) + 
  geom_histogram() + 
  facet_grid(rows = vars(redcap_event_name))
p1

```


EDSS 
```{r}
p2 <- ggplot(data = zv_fw_base_yr1_long, aes(x = demoEHR_EDSS_dateDiff)) + 
  geom_histogram() + 
  facet_grid(rows = vars(redcap_event_name))
p2
```
# Make Wide Dataset 

pivot wider 
```{r}
zv_fw_base_yr1_df_wide <- zv_fw_base_yr1_long %>% 
  select(-redcap_event_name)%>% 
  pivot_wider(id_cols = bw_id, 
              names_from = redcap_event_clean, 
              values_from = -c(bw_id, redcap_event_clean))
```

## Calculate and plot date diff from baseline to year 1 follow up 

## Calculate minimal significant change 
reference - https://www.sciencedirect.com/science/article/pii/S0966636220306998?ref=pdf_download&fr=RR-2&rr=940e64c029f9cba0 
Christopher bramah et al 
ICC3k: A fixed set of k judges rate each target. There is no generalization to a larger population of judges.
Think this means MDC is restricted to this data set? not sure 
```{r}
calc_srd95 <- function(data, SD) {
  
  # Calculate ICC
  icc_result <- ICC(data)
  print(icc_result)
  icc_result_df <- as.data.frame(icc_result$results, rownames = "icc_type")

  # Extract ICC3k 
  ICC_value <- icc_result_df$ICC[icc_result_df$type == "ICC3k"]
  print(ICC_value)
  
  # Calculate SEM - standard error of measurement 
  SEM <- SD * sqrt(1 - ICC_value)
  print(SEM)
  
  # Calculate SRD 
  SRD <- round(1.96 *  SEM * sqrt(2), 3)
  print(SRD)
  
  return(SRD)
}


```


delta pixel SRD95
```{r}
pixel_df <- zv_fw_base_yr1_df_wide %>% 
  select(delta_pix_h_rel_median_pose_zv_baseline, 
         delta_pix_h_rel_median_pose_zv_year2)

pixel_sd <- sd(c(pixel_df$delta_pix_h_rel_median_pose_zv_baseline, 
              pixel_df$delta_pix_h_rel_median_pose_zv_year2))

pixel_SRD  <- calc_srd95(pixel_df, pixel_sd)
pixel_srd_df <- data.frame(pixel_SRD_value = round(pixel_SRD, 2))
write.csv(pixel_srd_df,
          file.path(long_output_dir, 
                    'change_pixel_SRD.csv'))
```

stride time SRD95
```{r}
stride_time_df <-  zv_fw_base_yr1_df_wide %>% 
  select(stride_time_median_sec_pose_zv_baseline, 
         stride_time_median_sec_pose_zv_year2)

stride_time_sd <- sd(c(stride_time_df$stride_time_median_sec_pose_zv_baseline,
                       stride_time_df$stride_time_median_sec_pose_zv_year2))

stride_time_SRD <- calc_srd95(stride_time_df, stride_time_sd)
stride_time_srd_df <- data.frame(stride_time_SRD_value = round(stride_time_SRD, 2))
write.csv(stride_time_srd_df,
          file.path(long_output_dir, 
                    'change_stride_time_SRD.csv'))
```

## Visualize - paired boxplot from baseline to year2
```{r}
# Pixel Proxy 
ggplot(zv_fw_base_yr1_long, aes(x = redcap_event_name, y = delta_pix_h_rel_median_pose_zv)) +
  geom_boxplot(outlier.shape = NA, fill = "lightblue", width = 0.4) +
  geom_jitter(width = 0.1, aes(group = bw_id), alpha = 0.6, color = "black") +
  geom_line(aes(group = bw_id), color = "gray", alpha = 0.5) +
  theme_minimal() +
  labs(x = "Timepoint", y = "Pixel Proxy Median")


# Stride Time 
ggplot(zv_fw_base_yr1_long, aes(x = redcap_event_name, y = stride_time_median_sec_pose_zv)) +
  geom_boxplot(outlier.shape = NA, fill = "lightblue", width = 0.4) +
  geom_jitter(width = 0.1, aes(group = bw_id), alpha = 0.6, color = "black") +
  geom_line(aes(group = bw_id), color = "gray", alpha = 0.5) +
  theme_minimal() +
  labs(y = "Stride Time Median")


# T25FW 
ggplot(zv_fw_base_yr1_long, aes(x = redcap_event_name, y = clean_T25FW_Avg)) +
  geom_boxplot(outlier.shape = NA, fill = "lightblue", width = 0.4) +
  geom_jitter(width = 0.1, aes(group = bw_id), alpha = 0.6, color = "black") +
  geom_line(aes(group = bw_id), color = "gray", alpha = 0.5) +
  theme_minimal() +
  labs(y = "T25FW")


# EDSS
ggplot(zv_fw_base_yr1_long, aes(x = redcap_event_name, y = clean_EDSS)) +
  geom_boxplot(outlier.shape = NA, fill = "lightblue", width = 0.4) +
  geom_jitter(width = 0.1, aes(group = bw_id), alpha = 0.6, color = "black") +
  geom_line(aes(group = bw_id), color = "gray", alpha = 0.5) +
  theme_minimal() +
  labs(y = "EDSS")
```


## Calculate changes from baseline to year 2 
 
```{r}
tidy_wilcox <- function(test_object) {
  results_df <- data.frame(
    statistic = test_object$statistic,
    p_value = test_object$p.value,
    p_formatted = ifelse(test_object$p.value <0.001, '<0.001', round(test_object$p.value, 3)), 
    method = test_object$method,
    alternative = test_object$alternative
  ) %>% 
    mutate(p_formatted = as.character(p_formatted))
  
  
  return(results_df)
}
```

Year 2 values - baseline visit
```{r}
zv_fw_base_yr1_df_wide <- zv_fw_base_yr1_df_wide %>% 
  mutate(trialdate_diff_days = as.Date(trialdate_year2) - as.Date(trialdate_baseline),
         delta_pix_h_diff = delta_pix_h_rel_median_pose_zv_year2 - delta_pix_h_rel_median_pose_zv_baseline, 
         stride_time_diff = stride_time_median_sec_pose_zv_year2 - stride_time_median_sec_pose_zv_baseline, 
         t25fw_diff = clean_T25FW_Avg_year2 - clean_T25FW_Avg_baseline, 
         edss_diff = clean_EDSS_year2 - clean_EDSS_baseline) 

head(zv_fw_base_yr1_df_wide)
summary(zv_fw_base_yr1_df_wide)

```
### Date Diff 
```{r}
summary(as.numeric(zv_fw_base_yr1_df_wide$trialdate_diff_days))
hist(as.numeric(zv_fw_base_yr1_df_wide$trialdate_diff_days), probability = TRUE)
```

### Stats tests 
```{r}
# check the distribution of diff visually

hist(as.numeric(zv_fw_base_yr1_df_wide$delta_pix_h_diff), probability = TRUE)
hist(as.numeric(zv_fw_base_yr1_df_wide$stride_time_diff), probability = TRUE)
hist(as.numeric(zv_fw_base_yr1_df_wide$t25fw_diff), probability = TRUE)
hist(as.numeric(zv_fw_base_yr1_df_wide$edss_diff), probability = TRUE)

# Performs the Shapiro-Wilk test of normality
shapiro.test(as.numeric(zv_fw_base_yr1_df_wide$delta_pix_h_diff))
shapiro.test(as.numeric(zv_fw_base_yr1_df_wide$stride_time_diff))
shapiro.test(as.numeric(zv_fw_base_yr1_df_wide$t25fw_diff))
shapiro.test(as.numeric(zv_fw_base_yr1_df_wide$edss_diff))

# interpretation notes - delta pix h diff only normally distributed by shaprio test
# but... doesn't look normal on left end of plot, running wilcoxon signed rank test 
# for assumption not met for all variables 

# Performs a Wilcoxon Signed-Rank test if paired-data t-test assumption are not met
# delta pixel 
pixel_wcx_results <- wilcox.test(zv_fw_base_yr1_df_wide$delta_pix_h_rel_median_pose_zv_year2,
                                 zv_fw_base_yr1_df_wide$delta_pix_h_rel_median_pose_zv_baseline, 
                                 paired = TRUE,
                                 correct = FALSE, 
                                 exact = FALSE)

pixel_wcx_df <- tidy_wilcox(pixel_wcx_results) %>% 
  mutate(median_diff = round(median(zv_fw_base_yr1_df_wide$delta_pix_h_diff), 2), 
         q1 = round(quantile(zv_fw_base_yr1_df_wide$delta_pix_h_diff, probs = 0.25), 2),
         q3 = round(quantile(zv_fw_base_yr1_df_wide$delta_pix_h_diff, probs = 0.75), 2),
         data_name = pixel_wcx_results$data.name,
         name_clean = 'Pixel Height Proxy')

# stride time 
stride_wcx_results <- wilcox.test(zv_fw_base_yr1_df_wide$stride_time_median_sec_pose_zv_year2,
                                 zv_fw_base_yr1_df_wide$stride_time_median_sec_pose_zv_baseline, 
                                 paired = TRUE,
                                 correct = FALSE, 
                                 exact = FALSE)

stride_wcx_df <-tidy_wilcox(stride_wcx_results) %>%
  mutate(median_diff = round(median(zv_fw_base_yr1_df_wide$stride_time_diff), 2), 
         q1 = round(quantile(zv_fw_base_yr1_df_wide$stride_time_diff, probs = 0.25), 2),
         q3 = round(quantile(zv_fw_base_yr1_df_wide$stride_time_diff, probs = 0.75), 2), 
         data_name = stride_wcx_results$data.name, 
         name_clean = 'Stride Time')

# T25FW 
t25fw_wcx_results <- wilcox.test(zv_fw_base_yr1_df_wide$clean_T25FW_Avg_year2,
                                 zv_fw_base_yr1_df_wide$clean_T25FW_Avg_baseline, 
                                 paired = TRUE,
                                 correct = FALSE, 
                                 exact = FALSE)

t25fw_wcx_df <- tidy_wilcox(t25fw_wcx_results) %>%
  mutate(median_diff = round(median(zv_fw_base_yr1_df_wide$t25fw_diff), 2), 
         q1 = round(quantile(zv_fw_base_yr1_df_wide$t25fw_diff, probs = 0.25), 2),
         q3 = round(quantile(zv_fw_base_yr1_df_wide$t25fw_diff, probs = 0.75), 2),
         data_name = t25fw_wcx_results$data.name, 
         name_clean = 'T25FW') 

# EDSS 
edss_wcx_results <- wilcox.test(zv_fw_base_yr1_df_wide$clean_EDSS_year2,
                                 zv_fw_base_yr1_df_wide$clean_EDSS_baseline, 
                                 paired = TRUE,
                                 correct = FALSE, 
                                 exact = FALSE)

edss_wcx_df <- tidy_wilcox(edss_wcx_results) %>% 
  mutate(median_diff = round(median(zv_fw_base_yr1_df_wide$edss_diff), 2), 
         q1 = round(quantile(zv_fw_base_yr1_df_wide$edss_diff, probs = 0.25), 2),
         q3 = round(quantile(zv_fw_base_yr1_df_wide$edss_diff, probs = 0.75), 2),
         data_name = edss_wcx_results$data.name, 
         name_clean = 'EDSS')

```

```{r}
# bind and save all wilcoxong results 
wcx_df_all <- bind_rows(pixel_wcx_df, stride_wcx_df, t25fw_wcx_df, edss_wcx_df)
wcx_df_all
write.csv(wcx_df_all, 
          file.path(long_output_dir, 'fw_bw_tests_change_over_year.csv'))
```

 
## Add new columns for change 
Continuous change 
```{r}
 
zv_fw_base_yr1_df_wide <- zv_fw_base_yr1_df_wide %>% 
  mutate(delta_pix_change_cont = ifelse(delta_pix_h_diff == 0, 
                                         'No Change', 
                                         ifelse(delta_pix_h_diff > 0, 
                                                'Improved', 
                                                'Worsened')),
         stride_time_change_cont = ifelse(stride_time_diff == 0,
                                          'No Change', 
                                          ifelse(stride_time_diff > 0,
                                                 'Worsened',
                                                 'Improved')), 
         t25fw_change_cont = ifelse(t25fw_diff == 0, 
                                    'No Change', 
                                    ifelse(t25fw_diff > 0,
                                           'Worsened', 
                                           'Improved')), 
         edss_change_cont = ifelse(edss_diff == 0, 
                                   'No Change', 
                                   ifelse(edss_diff > 0,
                                          'Worsened',
                                          'Improved'))) 


```

Categorical Change 
If T25FW changes by more than 20% of baseline T25FW or video metrics change by more than SRD value 
== Change 

```{r}
zv_fw_base_yr1_df_wide <- zv_fw_base_yr1_df_wide %>% 
  mutate(delta_pix_change_cat = ifelse(abs(delta_pix_h_diff) < pixel_SRD, 
                                        'No Change',
                                        ifelse(delta_pix_h_diff > 0,
                                               'Improved', 
                                               'Worsened')), 
         stride_time_change_cat = ifelse(abs(stride_time_diff) < stride_time_SRD,
                                         'No Change', 
                                         ifelse(stride_time_diff > 0,
                                                'Worsened', 
                                                'Improved')), 
         t25fw_change_cat = ifelse(abs(t25fw_diff) < .2 * clean_T25FW_Avg_baseline, 
                                   'No Change', 
                                   ifelse(t25fw_diff > 0,
                                   'Worsened', 
                                   'Improved')),
         
         edss_change_cat = ifelse(clean_EDSS_baseline == 0 & edss_diff >= 1.5,
                                  'Worsened', 
                                  ifelse(clean_EDSS_baseline >= 1.0 & clean_EDSS_baseline <=5.0 & edss_diff >= 1.0,
                                         'Worsened', 
                                         ifelse(clean_EDSS_baseline >= 5.5 & edss_diff >= 0.5,
                                                'Worsened',
                                                'Other')))
         )
                                           
```

Binary - worsened yes no 

```{r}
zv_fw_base_yr1_df_wide <- zv_fw_base_yr1_df_wide %>% 
  mutate(delta_pix_worsenYN_cont = ifelse(delta_pix_change_cont == 'Worsened', 
                                           'Yes', 'No'), 
         stride_time_worsenYN_cont = ifelse(stride_time_change_cont == 'Worsened',
                                        'Yes', 'No'), 
         either_pix_stride_worsenYN_cont = ifelse((delta_pix_change_cont == 'Worsened')|
                                                   (stride_time_change_cont == 'Worsened'), 
                                                  'Yes', 'No'),
         t25fw_worsenYN_cont = ifelse(t25fw_change_cont == 'Worsened',
                                      'Yes', 'No'), 
         edss_worsenYN_cont = ifelse(edss_change_cont == 'Worsened',
                                     'Yes', 'No'), 
         delta_pix_worsenYN_cat = ifelse(delta_pix_change_cat == 'Worsened',
                                         'Yes', 'No'), 
         stride_time_worsenYN_cat = ifelse(stride_time_change_cat == 'Worsened',
                                           'Yes', 'No'), 
         either_pix_stride_worsenYN_cat = ifelse((delta_pix_change_cat == 'Worsened') | 
                                                  (stride_time_change_cat == 'Worsened'), 
                                                 'Yes', 'No'), 
         t25fw_worsenYN_cat = ifelse(t25fw_change_cat == 'Worsened',
                                     'Yes', 'No'), 
         edss_worsenYN_cat = ifelse(edss_change_cat == 'Worsened', 
                                    'Yes', 'No'))

write.csv(zv_fw_base_yr1_df_wide,
          file.path(long_output_dir, 
                    'fw_base_yer2_wide_groups.csv')) 
         
         
```


```{r}
print('t25fw_worsenYN_cont')
print(table(zv_fw_base_yr1_df_wide$t25fw_worsenYN_cont))

print('t25fw_worsenYN_cat')
print(table(zv_fw_base_yr1_df_wide$t25fw_worsenYN_cat))
```

```{r}
print('edss_worsenYN_cont')
print(table(zv_fw_base_yr1_df_wide$edss_worsenYN_cont))

print('edss_worsenYN_cat')
print(table(zv_fw_base_yr1_df_wide$edss_worsenYN_cat))
```

```{r}
print('delta_pix_worsenYN_cont')
print(table(zv_fw_base_yr1_df_wide$delta_pix_worsenYN_cont))

print('delta_pix_worsenYN_cat')
print(table(zv_fw_base_yr1_df_wide$delta_pix_worsenYN_cat))

```
```{r}
print('stride_time_worsenYN_cont')
print(table(zv_fw_base_yr1_df_wide$stride_time_worsenYN_cont))

print('stride_time_worsenYN_cat')
print(table(zv_fw_base_yr1_df_wide$stride_time_worsenYN_cat))

```
```{r}
print('either video metric cont')
print(table(zv_fw_base_yr1_df_wide$either_pix_stride_worsenYN_cont))

print('either video metric cat')
print(table(zv_fw_base_yr1_df_wide$either_pix_stride_worsenYN_cat))
```

## Create new column for following: 
Worsened in both video and clinical = 1
Worsened clinical measure only = 2
Worsened video measure only = 3
Worsened in neither video or clinical = 4

column name = videometric_vs_clinicalmetric_datatype_groups 
```{r}
add_video_vs_clinical_groups <- function(df, video_col, clinical_col, new_col_name) {
  df[[new_col_name]] <- with(df, ifelse(df[[video_col]] == "Yes" & df[[clinical_col]] == "Yes", 'Both', # both 
                                 ifelse(df[[video_col]] == "No" & df[[clinical_col]] == "Yes", 'Clinical Only', # clinical only
                                 ifelse(df[[video_col]] == "Yes"  & df[[clinical_col]] == "No", 'Video Only', # video only 
                                                                    'Neither')))) # neither 
  df[[new_col_name]] <- as.factor(df[[new_col_name]])
  return(df)
}

```

### T25FW Continuous vs Video Continuous 
```{r}
# delta pixel 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'delta_pix_worsenYN_cont',
                                                       't25fw_worsenYN_cont',
                                                       't25fw_cont_vs_pixel_cont_groups')

table(zv_fw_base_yr1_df_wide$t25fw_cont_vs_pixel_cont_groups)

```

```{r}
# stride time 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'stride_time_worsenYN_cont',
                                                       't25fw_worsenYN_cont',
                                                       't25fw_cont_vs_stride_cont_groups')

table(zv_fw_base_yr1_df_wide$t25fw_cont_vs_stride_cont_groups)
```

```{r}
# either pixel or stride time 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'either_pix_stride_worsenYN_cont',
                                                       't25fw_worsenYN_cont',
                                                       't25fw_cont_vs_either_vid_cont_groups')

table(zv_fw_base_yr1_df_wide$t25fw_cont_vs_either_vid_cont_groups)
```


### T25FW Categorical vs Video Categorical 
```{r}
# Delta pixel 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'delta_pix_worsenYN_cat',
                                                       't25fw_worsenYN_cat',
                                                       't25fw_cat_vs_pixel_cat_groups')

table(zv_fw_base_yr1_df_wide$t25fw_cat_vs_pixel_cat_groups)

```
```{r}
# Stride Time 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'stride_time_worsenYN_cat',
                                                       't25fw_worsenYN_cat',
                                                       't25fw_cat_vs_stride_cat_groups')

table(zv_fw_base_yr1_df_wide$t25fw_cat_vs_stride_cat_groups)
```
```{r}
# either pixel or stride time 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'either_pix_stride_worsenYN_cat',
                                                       't25fw_worsenYN_cat',
                                                       't25fw_cat_vs_either_vid_cat_groups')

table(zv_fw_base_yr1_df_wide$t25fw_cat_vs_either_vid_cat_groups)
```

### EDSS continuous vs Video Continuous 
```{r}
# pixel 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'delta_pix_worsenYN_cont',
                                                       'edss_worsenYN_cont',
                                                       'edss_cont_vs_pixel_cont_groups')

table(zv_fw_base_yr1_df_wide$edss_cont_vs_pixel_cont_groups)

```
```{r}
# stride time 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'stride_time_worsenYN_cont',
                                                       'edss_worsenYN_cont',
                                                       'edss_cont_vs_stride_cont_groups')

table(zv_fw_base_yr1_df_wide$edss_cont_vs_stride_cont_groups)
```
```{r}
# either pixel or stride time 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'either_pix_stride_worsenYN_cont',
                                                       'edss_worsenYN_cont',
                                                       'edss_cont_vs_either_vid_cont_groups')

table(zv_fw_base_yr1_df_wide$edss_cont_vs_either_vid_cont_groups)
```


### EDSS categorical vs video categorical 
```{r}
# pixel 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'delta_pix_worsenYN_cat',
                                                       'edss_worsenYN_cat',
                                                       'edss_cat_vs_pixel_cat_groups')

table(zv_fw_base_yr1_df_wide$edss_cat_vs_pixel_cat_groups)
```

```{r}
# stride time 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'stride_time_worsenYN_cat',
                                                       'edss_worsenYN_cat',
                                                       'edss_cat_vs_stride_cat_groups')

table(zv_fw_base_yr1_df_wide$edss_cat_vs_pixel_cat_groups)
```
```{r}
# either pixel or stride time 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'either_pix_stride_worsenYN_cat',
                                                       'edss_worsenYN_cat',
                                                       'edss_cat_vs_either_vid_cat_groups')

table(zv_fw_base_yr1_df_wide$edss_cat_vs_either_vid_cat_groups)
```
### T25FW categorical vs video continuous 
```{r}
# delta pixel 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'delta_pix_worsenYN_cont',
                                                       't25fw_worsenYN_cat',
                                                       't25fw_cat_vs_pixel_cont_groups')

table(zv_fw_base_yr1_df_wide$t25fw_cat_vs_pixel_cont_groups)
```
```{r}
# stride time  
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'stride_time_worsenYN_cont',
                                                       't25fw_worsenYN_cat',
                                                       't25fw_cat_vs_stride_cont_groups')

table(zv_fw_base_yr1_df_wide$t25fw_cat_vs_stride_cont_groups)
```

```{r}
# either 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'either_pix_stride_worsenYN_cont',
                                                       't25fw_worsenYN_cat',
                                                       't25fw_cat_vs_either_vid_cont_groups')

table(zv_fw_base_yr1_df_wide$t25fw_cat_vs_either_vid_cont_groups)

```

### EDSS categorical vs video metrics 
```{r}
# pixel 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'delta_pix_worsenYN_cont',
                                                       'edss_worsenYN_cat',
                                                       'edss_cat_vs_pixel_cont_groups')

table(zv_fw_base_yr1_df_wide$edss_cat_vs_pixel_cont_groups)
```


```{r}
# stride time 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'stride_time_worsenYN_cont',
                                                       'edss_worsenYN_cat',
                                                       'edss_cat_vs_stride_cont_groups')

table(zv_fw_base_yr1_df_wide$edss_cat_vs_stride_cont_groups)
```

```{r}
# either 
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'either_pix_stride_worsenYN_cont',
                                                       'edss_worsenYN_cat',
                                                       'edss_cat_vs_either_vid_cont_groups')

table(zv_fw_base_yr1_df_wide$edss_cat_vs_either_vid_cont_groups)
```



## Plot all 

```{r}
# convert to long dataframe 
plot_long_df <- zv_fw_base_yr1_df_wide %>% 
  select(c(bw_id, ends_with('groups'))) %>% 
  pivot_longer(cols = ends_with('groups'), 
               names_to = 'cols_compared', 
               values_to = 'agreement_group') %>% 
  mutate(clinical_outcome = ifelse(grepl("t25fw_cont", cols_compared), 
                               'T25FW Continuous', 
                               ifelse(grepl("t25fw_cat", cols_compared),
                                      'T25FW Categorical', 
                                      ifelse(grepl('edss_cont', cols_compared),
                                             'EDSS Ordinal',
                                             ifelse(grepl('edss_cat', cols_compared),
                                                    'EDSS Categorical', 
                                                    'Other'))))) %>%   # add column for outcome 
  mutate(video_metric = ifelse(grepl('pixel_cont', cols_compared), 
                               'Pixel Proxy Continuous', 
                               ifelse(grepl('pixel_cat', cols_compared), 
                                      'Pixel Proxy Categorical', 
                                      ifelse(grepl('stride_cont', cols_compared),
                                             'Stride Time Continuous',
                                             ifelse(grepl('stride_cat', cols_compared),
                                                    'Stride Time Categorical',
                                                    ifelse(grepl('either_vid_cont', cols_compared),
                                                           'Either Continuous',
                                                           ifelse(grepl('either_vid_cat', cols_compared),
                                                                  'Either Categorical',
                                                                  'Other'))))))) # add column for video metric 



# factor: agreement_group
plot_long_df$agreement_group <- factor(plot_long_df$agreement_group, 
                                       levels = c('Video Only', 
                                                  'Clinical Only', 
                                                  'Both', 
                                                  'Neither'), 
                                       labels = c('Video metric only', 
                                                  'Clinical outcome only', 
                                                  'Both',
                                                  'Neither'))

# Plot all 
ggplot(data = plot_long_df, aes(x = video_metric, fill = agreement_group)) + 
  geom_bar(position = 'fill') + 
  facet_grid(rows = vars(clinical_outcome)) +
   scale_fill_brewer(name = 'Worsened by:', palette = "Set2") +
  labs(y = "Proportion", x = "") +
  theme_minimal() + 
  theme(legend.position = 'bottom', 
        axis.text.x = element_text(angle = 90), 
        strip.text.y.right = element_text(angle = 0))
```
# Attempt 1: Agreement between videos and clinical measures: Increased/Decrease/No Change
Figure legend 
left: EDSS vs continuous video metrics 
middle: continuous25FW vs continuous video metrics 
left: categorical T25FW (change by 20% or more) and video metric categorical (change of SRD95 or more)

left vs middle: both continuous video metrics, higher % showing change in video metric only in EDSS. Continuous video metrics more sensitive to change than EDSS (% with decrease in . T25FW continous 

Right:categorical - strong agreement between videos and T25Fw - % both and neither in these groups 

```{r}
exclude_strings <- c('edss_cont_vs_pixel_cat_groups',
                     'edss_cont_vs_stride_cat_groups',
                     "t25fw_cont_vs_pixel_cat_groups",
                     't25fw_cont_vs_stride_cat_groups'
                     )

plot_long_df_2 <- plot_long_df %>%  
  filter(!grepl(paste(exclude_strings, collapse="|"), cols_compared))

# make new columns 
# only comparing stride time categorical and pixel categorical to T25FW categorical 
# same for continuou 
# don't need to specificy categorical or continuous on plot 
plot_long_df_2 <- plot_long_df_2 %>% 
  mutate(video_metric_simple = ifelse(grepl('Pixel Proxy', video_metric),
                                      'Pixel Proxy', 
                                      ifelse(grepl('Stride Time', video_metric),
                                             'Stride Time', 
                                             ifelse(grepl('Either', video_metric), 
                                                    'Either', 'Other'))))
plot_long_df_2$video_metric_simple = factor(plot_long_df_2$video_metric_simple, 
                                            levels = c('Pixel Proxy', 
                                                       'Stride Time', 
                                                       'Either'))

print(table(plot_long_df_2$clinical_outcome, plot_long_df_2$video_metric)) # confirm

# factor clinical outcome variable - control order of facet_wrap
plot_long_df_2$clinical_outcome <- factor(plot_long_df_2$clinical_outcome, 
                                          levels = c('T25FW Continuous', 
                                                     'T25FW Categorical',
                                                     'EDSS Ordinal', 
                                                     'EDSS Categorical'))
```

Plot wihtout numbers 
```{r}
ggplot(data = plot_long_df_2, aes(x = video_metric_simple, fill = agreement_group)) + 
  geom_bar(position = 'fill') + 
  facet_wrap(~clinical_outcome) +
   scale_fill_brewer(name = 'Worsened by:', palette = "Set2") +
  labs(y = "Proportion", x = "Video Metric") +
  theme_minimal() + 
  theme(legend.position = 'bottom')

```
## Plot with proportion
```{r}
# add proportion column to put text on plot 
plot_long_df_2_prop <- plot_long_df_2 %>%
  group_by(clinical_outcome, video_metric_simple, agreement_group) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(clinical_outcome, video_metric_simple) %>%
  mutate(prop = n / sum(n),
         label = scales::percent(prop, accuracy = 1))

# proportion 
p <- ggplot(plot_long_df_2_prop, aes(x = video_metric_simple, y = prop, fill = agreement_group)) +
  geom_bar(stat = "identity", position = "fill", width = 0.5) +
  geom_text(aes(label = ifelse(prop > 0.03, label, "")),  # Only label if > 5%
            position = position_fill(vjust = 0.5),
            size = 3.25) + 
  facet_wrap(~ clinical_outcome) +
   scale_fill_brewer(name = "Worsened by:", palette = "Set2") +
  labs(y = "Proportion", x = "") +
  theme_minimal() + 
  theme(plot.title = element_text(size = 9, face = 'bold', hjust = 0.5),
          axis.title = element_text(size = 9),
          #axis.text = element_text(size = 9),
          axis.text.x = element_text(size = 9, face = 'italic'),
          axis.text.y = element_blank(),
          axis.title.y = element_blank(),
          axis.ticks.y = element_blank(),
          legend.text = element_text(size = 9), 
          legend.title = element_text(size = 9),
          legend.position = 'bottom',
          legend.key.size = unit(0.5, "lines"))  
p
ggsave(file.path(long_output_dir, 'barcharts_proportion_1.png'),
       bg = 'white', height = 5, width = 5)

```

# Attempt 2: Only categorical change in outcomes 
Of people with categorical change by T25FW or EDSS, are video metrics absolute change trending in correct direction? 

```{r}
# check - agreement between EDSS and T25FW categorical change?
head(zv_fw_base_yr1_df_wide)
```

```{r}
#  check - agreement between EDSS and T25FW categorical change? 
# do participants worsen in both?

table(zv_fw_base_yr1_df_wide$t25fw_change_cat)

table(zv_fw_base_yr1_df_wide$edss_change_cat)

# worsened by both edss and t25FW 
both_df <- zv_fw_base_yr1_df_wide %>% 
  filter(t25fw_change_cat == 'Worsened' & edss_change_cat == 'Worsened')
both_df

```

Worsened by EDSS categorical
```{r}
edss_df <- zv_fw_base_yr1_df_wide %>% 
  filter(edss_change_cat == 'Worsened') 

edss_df %>% select(bw_id, clean_EDSS_baseline, 
                   clean_EDSS_year2, 
                   delta_pix_change_cont, 
                   stride_time_change_cont)

# categorical worsening by EDSS only for people with EDSS < 4 at both visits - expect much affect on walking?
```

Worsened by T25FW 

```{r}
# T25FW categorical 
t25fw_df <- zv_fw_base_yr1_df_wide %>% 
  filter(t25fw_change_cat == 'Worsened')
  
t25fw_df %>% select(bw_id, clean_T25FW_Avg_baseline, 
                    clean_T25FW_Avg_year2, 
                    delta_pix_change_cont, 
                    stride_time_change_cont)

```
Change in pixel height 
```{r}
table(zv_fw_base_yr1_df_wide$delta_pix_change_cont,
      zv_fw_base_yr1_df_wide$t25fw_change_cat)


table(zv_fw_base_yr1_df_wide$delta_pix_worsenYN_cont,
      zv_fw_base_yr1_df_wide$t25fw_worsenYN_cat)

```
```{r}
table(zv_fw_base_yr1_df_wide$stride_time_change_cont,
      zv_fw_base_yr1_df_wide$t25fw_change_cat)

table(zv_fw_base_yr1_df_wide$stride_time_worsenYN_cont,
      zv_fw_base_yr1_df_wide$t25fw_worsenYN_cat)
```
```{r}
table(zv_fw_base_yr1_df_wide$delta_pix_change_cont,
      zv_fw_base_yr1_df_wide$edss_change_cat)


table(zv_fw_base_yr1_df_wide$delta_pix_worsenYN_cont,
      zv_fw_base_yr1_df_wide$edss_worsenYN_cat)

```
```{r}
table(zv_fw_base_yr1_df_wide$stride_time_change_cont,
      zv_fw_base_yr1_df_wide$edss_change_cat)

table(zv_fw_base_yr1_df_wide$stride_time_worsenYN_cont,
      zv_fw_base_yr1_df_wide$edss_worsenYN_cat)
```



## Format for plotting  
```{r}
include_strings <- c('edss_cat_vs_either_vid_cont_groups',
                     'edss_cat_vs_pixel_cont_groups',
                     "edss_cat_vs_stride_cont_groups", 
                     't25fw_cat_vs_either_vid_cont_groups',
                     't25fw_cat_vs_pixel_cont_groups',
                     "t25fw_cat_vs_stride_cont_groups" 
                     )

plot_long_df_3 <- plot_long_df %>%  
  filter(grepl(paste(include_strings, collapse="|"), cols_compared))  

table(plot_long_df_3$cols_compared)   
```

## Plot with proportion 
```{r}
# add proportion column to put text on plot 
plot_long_df_3_prop <- plot_long_df_3 %>%
  group_by(clinical_outcome, video_metric, agreement_group) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(clinical_outcome, video_metric) %>%
  mutate(prop = n / sum(n),
         label = scales::percent(prop, accuracy = 1))

# proportion 
p <- ggplot(plot_long_df_3_prop, aes(x = video_metric, y = prop, fill = agreement_group)) +
  geom_bar(stat = "identity", position = "fill", width = 0.5) +
  geom_text(aes(label = ifelse(prop > 0.03, label, "")),  # Only label if > 5%
            position = position_fill(vjust = 0.5),
            size = 3.25) + 
  facet_wrap(~ clinical_outcome) +
   scale_fill_brewer(name = "Worsened by:", palette = "Set2") +
  labs(y = "Proportion", x = "") +
  theme_minimal() + 
  theme(plot.title = element_text(size = 9, face = 'bold', hjust = 0.5),
          axis.title = element_text(size = 9),
          #axis.text = element_text(size = 9),
          axis.text.x = element_text(size = 9, face = 'italic', angle = 90),
          axis.text.y = element_blank(),
          axis.title.y = element_blank(),
          axis.ticks.y = element_blank(),
          legend.text = element_text(size = 9), 
          legend.title = element_text(size = 9),
          legend.position = 'bottom',
          legend.key.size = unit(0.5, "lines"))  
p
ggsave(file.path(long_output_dir, 'barcharts_proportion_2.png'),
       bg = 'white', height = 5, width = 5)
```


# Attempt 3: PCA 
## Principal component 
principal component of stride time and pixel from all videos from all participants 
```{r}
# load all FW data 
fw_clean_data_path <- file.path("C:/Users/mmccu/Box/MM_Personal/5_Projects/BoveLab/3_Data_and_Code/gait_bw_zeno_home_analysis",
                       analysis_version, 
                       "000_merged_cleaned_data", 
                       "zv_bw_merged_gait_vertical_FW_1_clean.csv")

zv_fw_clean_all_df <- read.csv(fw_clean_data_path)

```

filter just participants with both stride time and pixel proxy 
```{r}
zv_fw_clean_df <- zv_fw_clean_all_df %>% 
  drop_na(c("stride_time_median_sec_pose_zv", 
            "delta_pix_h_rel_median_pose_zv",
            "mean_cadence_step_per_min_pose_zv",
            "stride_width_median_cm_pose_zv"))
           
            
            
             #"stance_time_sec_std_pose_zv", 
           # "swing_time_sec_std_pose_zv"))
  
  
 # filter(!is.na(delta_pix_h_rel_median_pose_zv) & !is.na(stride_time_median_sec_pose_zv))

nrow(zv_fw_clean_df)
```

log transform delta pixel proxy 
```{r}
plot(zv_fw_clean_df$delta_pix_h_rel_median_pose_zv,
     zv_fw_clean_df$clean_T25FW_Avg)


plot(log(zv_fw_clean_df$delta_pix_h_rel_median_pose_zv),
     log(zv_fw_clean_df$clean_T25FW_Avg))

zv_fw_clean_df$log_delta_pixel_median <- log(zv_fw_clean_df$delta_pix_h_rel_median_pose_zv)
```

principal component 
```{r}
pca_pix_stride <- prcomp(~ log_delta_pixel_median + 
                           stride_time_median_sec_pose_zv, 
                         data = zv_fw_clean_df, 
                         scale = TRUE)

summary(pca_pix_stride)
pca_pix_stride$rotation 
# PC1 - faster change in pixel height and shorter stride times (think higher value in this = faster walking??); decreased pc1 = worsening in walking

biplot(pca_pix_stride)
```







## Cross sectional 
Add PCA scores to data frame 

```{r}
pca_scores <- pca_pix_stride$x

zv_fw_clean_wPCs <- cbind(zv_fw_clean_df, pca_scores) 
head(zv_fw_clean_wPCs)
write.csv(zv_fw_clean_wPCs, file.path(pca_output_dir,
                                      'zv_fw_wPC_scores.csv'))
```

### plot T25FW vs PC1 
PC1 - caculated for all participants 
Plots - just people with baseline and follow up, both visits plotted 
```{r}
hist(zv_fw_clean_wPCs$clean_T25FW_Avg)


hist(zv_fw_clean_wPCs$PC1)

# original T25FW data 
plot(zv_fw_clean_wPCs$clean_T25FW_Avg,
     zv_fw_clean_wPCs$PC1)

cor.test(zv_fw_clean_wPCs$clean_T25FW_Avg, 
         zv_fw_clean_wPCs$PC1, 
         method=c("spearman")) # non-parametric 

cor.test(zv_fw_clean_wPCs$clean_T25FW_Avg, 
         zv_fw_clean_wPCs$delta_pix_h_rel_median_pose_zv, 
         method=c("spearman")) # non-parametric 


cor.test(zv_fw_clean_wPCs$clean_T25FW_Avg, 
         zv_fw_clean_wPCs$stride_time_median_sec_pose_zv, 
         method=c("spearman")) # non-parametric 


# Log transform T25FW 
zv_fw_clean_wPCs$log_t25fw <- log(zv_fw_clean_wPCs$clean_T25FW_Avg)

plot(zv_fw_clean_wPCs$log_t25fw,
     zv_fw_clean_wPCs$PC1)

hist(zv_fw_clean_wPCs$log_t25fw)

# spearman corr will be the same, based on rank 
```
### mixed effects regression 
pixel and stride time principal component 
```{r}
library(lme4)
library(lmerTest)
library(performance)

# original T25FW  
lme_results <- lmer(clean_T25FW_Avg ~ PC1 + clean_Age + (1|bw_id), 
                    data = zv_fw_clean_wPCs,
                    REML = FALSE)

summary(lme_results)
hist(residuals(lme_results))
r2(lme_results)$R2_conditional
r2(lme_results)$R2_marginal

# log T25FW + age 
lme_log_results <- lmer(log_t25fw ~ PC1 + clean_Age 
                        + (1|bw_id), 
                        data = zv_fw_clean_wPCs,
                        REML = FALSE)

summary(lme_log_results)
hist(residuals(lme_log_results))
r2(lme_log_results)$R2_conditional
r2(lme_log_results)$R2_marginal

# save results 
lme_results_df <- tidy(lme_log_results, conf.int = TRUE) %>% 
    mutate(R2_conditional = r2(lme_log_results)$R2_conditional) %>% 
    mutate(R2_marginal = r2(lme_log_results)$R2_marginal) %>% 
    mutate(Outcome = 'log_t25fw') %>% 
    mutate(p_val_formatted = ifelse(p.value < 0.001, '<0.001', round(p.value, 3)))
lme_results_df
write.csv(lme_results_df, file.path(pca_output_dir,
                                    'pc1_mixed_effect_results.csv'))
```


### Try pixel, stride, and cadence pca 

```{r}
# pixel stride time cadence

pca_pix_stride_cadence <- prcomp(~ log_delta_pixel_median + 
                           stride_time_median_sec_pose_zv + 
                           mean_cadence_step_per_min_pose_zv, 
                         data = zv_fw_clean_df, 
                         scale = TRUE)

summary(pca_pix_stride_cadence)
pca_pix_stride_cadence$rotation 

biplot(pca_pix_stride_cadence)

# add to df 
pca_scores_test <- pca_pix_stride_cadence$x

zv_fw_clean_wPCs_test <- cbind(zv_fw_clean_df, pca_scores_test) 
head(zv_fw_clean_wPCs_test)

## plot vs T25FW 
# original T25FW data 
plot(zv_fw_clean_wPCs_test$clean_T25FW_Avg,
     zv_fw_clean_wPCs_test$PC1)

cor.test(zv_fw_clean_wPCs_test$clean_T25FW_Avg, 
         zv_fw_clean_wPCs_test$PC1, 
         method=c("spearman")) # non-parametric 

zv_fw_clean_wPCs_test$log_t25fw <- log(zv_fw_clean_wPCs_test$clean_T25FW_Avg)

plot(zv_fw_clean_wPCs_test$log_t25fw,
     zv_fw_clean_wPCs_test$PC1)

# regression 
# log T25FW + age 
lme_log_results_test <- lmer(log_t25fw ~ PC1 + clean_Age 
                        + (1|bw_id), 
                        data = zv_fw_clean_wPCs_test,
                        REML = FALSE)

summary(lme_log_results_test)
hist(residuals(lme_log_results_test))
r2(lme_log_results_test)$R2_conditional
r2(lme_log_results_test)$R2_marginal
```


### Plots 
```{r}
# observed vs predicted 
predicted <- predict(lme_log_results)
df <- data.frame(Observed = zv_fw_clean_wPCs$log_t25fw,
	                 Predicted = predicted)

p1 <- ggplot(df, aes(x = Observed, y = Predicted)) +
	 geom_point(size = 0.75, alpha = 0.6) +
	 geom_abline(intercept = 3, slope = 1,
	             linetype = "dashed") + 
  coord_fixed(ratio = 1) + 
  labs(title='PC1 + Age: Observed vs Predicted',
        x ="Observed Log T25FW", y = "Predicted Log T25FW") + 
  theme_minimal() + 
  theme(plot.title = element_text(size = 9, face = 'bold', hjust = 0.5),
          axis.title = element_text(size = 9),
          #axis.text = element_text(size = 9),
          axis.text.y = element_text(size = 9),
          axis.text.x = element_text(size = 9),
          legend.text = element_text(size = 9))  
p1
ggsave(file.path(pca_output_dir,
                 'pc1_mixed_effect_obs_vs_pred.png'),
       bg = 'white', height = 3, width = 3)



# fitted vs residuals 
fitted_vals <- fitted(lme_log_results)
resid_vals <- resid(lme_log_results)

resid_df <- data.frame(Fitted = fitted_vals, Residuals = resid_vals)

p2 <- ggplot(resid_df, aes(x = Fitted, y = Residuals)) +
  geom_point(alpha = 0.6, size = 0.75) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey") +
  labs(
    title = paste('PC1 + Age: Fitted vs Residuals'),
    x = "Fitted Values",
    y = "Residuals"
  ) +
  coord_fixed(ratio = 1) +
  theme_minimal() + 
  theme(plot.title = element_text(size = 9, face = 'bold', hjust = 0.5),
          axis.title = element_text(size = 9),
          #axis.text = element_text(size = 9),
          axis.text.y = element_text(size = 9),
          axis.text.x = element_text(size = 9),
          legend.text = element_text(size = 9))  
p2
ggsave(file.path(pca_output_dir,
                 'pc1_mixed_effect_fitted_vs_residual.png'), 
       bg = 'white', height = 3, width = 3)

```


## Longitudinal 
```{r}
# to-do: rename clean wPC from here down 

longitud_zv_fw_clean_wPCs <- zv_fw_clean_wPCs %>% 
  filter(id_date_pose_zv %in% zv_fw_base_yr1_long$id_date_pose_zv) %>% # only select participants with base + 1yr vists 
  mutate(redcap_event_clean = ifelse(redcap_event_name == 'Brainwalk: Baseline visit (Arm 1: Baseline visit)',
                                     'baseline',
                                     'year2')) ## make clean BW visit 
head(longitud_zv_fw_clean_wPCs) 
```

Pivot wider 
```{r}
zv_fw_clean_wPCs_wide <- longitud_zv_fw_clean_wPCs %>% 
  select(bw_id, redcap_event_clean, id_date_pose_zv, PC1, PC2) %>% # select just id, redcap event, and PCs 
  pivot_wider(id_cols = bw_id, 
              names_from = redcap_event_clean, 
              values_from = -c(bw_id, redcap_event_clean)) 

head(zv_fw_clean_wPCs_wide)
```

merge PCs with existing wide data frame 
```{r}
zv_fw_base_yr1_df_wide <- merge(zv_fw_base_yr1_df_wide, zv_fw_clean_wPCs_wide, 
                                by = 'bw_id')
view(zv_fw_base_yr1_df_wide)
```

Change in PC1 
```{r}
zv_fw_base_yr1_df_wide <- zv_fw_base_yr1_df_wide %>% 
  mutate(pc1_diff = PC1_year2 - PC1_baseline,
         pc2_diff = PC2_year2 - PC2_baseline) %>%  # add columns with change from baseline to follow up 
  mutate(pc1_worsen = ifelse(pc1_diff > 0, 'No', 'Yes'))  
  # from loading - increase in PC1 = better walking (greater change in pixel height and quicker strides)
  

```

### tables 
T25FW 
```{r}
table(zv_fw_base_yr1_df_wide$pc1_worsen,
                 zv_fw_base_yr1_df_wide$t25fw_worsenYN_cat)


table(zv_fw_base_yr1_df_wide$pc1_worsen,
                 zv_fw_base_yr1_df_wide$t25fw_worsenYN_cont)

```
EDSS 
```{r}
table(zv_fw_base_yr1_df_wide$pc1_worsen,
                 zv_fw_base_yr1_df_wide$edss_worsenYN_cat)


table(zv_fw_base_yr1_df_wide$pc1_worsen,
                 zv_fw_base_yr1_df_wide$edss_worsenYN_cont)
```


### new columns 
T25FW continuous 
```{r}
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'pc1_worsen',
                                                       't25fw_worsenYN_cont',
                                                       't25fw_cont_vs_pc1_groups')

table(zv_fw_base_yr1_df_wide$t25fw_cont_vs_pc1_groups)
```
T25FW categorical 

```{r}
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'pc1_worsen',
                                                       't25fw_worsenYN_cat',
                                                       't25fw_cat_vs_pc1_groups')

table(zv_fw_base_yr1_df_wide$t25fw_cat_vs_pc1_groups)
```
EDSS continuous 
```{r}
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'pc1_worsen',
                                                       'edss_worsenYN_cont',
                                                       'edss_cont_vs_pc1_groups')

table(zv_fw_base_yr1_df_wide$edss_cont_vs_pc1_groups)
```
EDSS Categorical 
```{r}
zv_fw_base_yr1_df_wide <- add_video_vs_clinical_groups(zv_fw_base_yr1_df_wide, 
                                                       'pc1_worsen',
                                                       'edss_worsenYN_cat',
                                                       'edss_cat_vs_pc1_groups')

table(zv_fw_base_yr1_df_wide$edss_cat_vs_pc1_groups)
write.csv(zv_fw_base_yr1_df_wide, file.path(long_output_dir,
                                            'fw_base_year2_wide_groups_wPCs.csv'))
```

### Format for plotting 
```{r}
# convert to long dataframe 
plot_long_df_pca <- zv_fw_base_yr1_df_wide %>% 
  select(c(bw_id, ends_with('groups'))) %>% 
  pivot_longer(cols = ends_with('groups'), 
               names_to = 'cols_compared', 
               values_to = 'agreement_group') %>% 
  mutate(clinical_outcome = ifelse(grepl("t25fw_cont", cols_compared), 
                               'T25FW Continuous', 
                               ifelse(grepl("t25fw_cat", cols_compared),
                                      'T25FW Categorical', 
                                      ifelse(grepl('edss_cont', cols_compared),
                                             'EDSS Ordinal',
                                             ifelse(grepl('edss_cat', cols_compared),
                                                    'EDSS Categorical', 
                                                    'Other'))))) %>%   # add column for outcome 
  mutate(video_metric = ifelse(grepl('pixel_cont', cols_compared), 
                               'Pixel Proxy Continuous', 
                               ifelse(grepl('pixel_cat', cols_compared), 
                                      'Pixel Proxy Categorical', 
                                      ifelse(grepl('stride_cont', cols_compared),
                                             'Stride Time Continuous',
                                             ifelse(grepl('stride_cat', cols_compared),
                                                    'Stride Time Categorical',
                                                    ifelse(grepl('either_vid_cont', cols_compared),
                                                           'Either Continuous',
                                                           ifelse(grepl('either_vid_cat', cols_compared),
                                                                  'Either Categorical',
                                                                  ifelse(grepl('pc1', cols_compared), 
                                                                               'PC1',
                                                                               'Other')))))))) # add column for video metric 



# factor: agreement_group
plot_long_df_pca$agreement_group <- factor(plot_long_df_pca$agreement_group, 
                                       levels = c('Video Only', 
                                                  'Clinical Only', 
                                                  'Both', 
                                                  'Neither'), 
                                       labels = c('Video metric only', 
                                                  'Clinical outcome only', 
                                                  'Both',
                                                  'Neither'))

# Plot all 
ggplot(data = plot_long_df_pca, aes(x = video_metric, fill = agreement_group)) + 
  geom_bar(position = 'fill') + 
  facet_grid(rows = vars(clinical_outcome)) +
   scale_fill_brewer(name = 'Worsened by:', palette = "Set2") +
  labs(y = "Proportion", x = "") +
  theme_minimal() + 
  theme(legend.position = 'bottom', 
        axis.text.x = element_text(angle = 90), 
        strip.text.y.right = element_text(angle = 0))
```

Filter to only include: 
  PC1 comparison 
  continuous vs continuous 
  categorical vs categorical 
```{r}
# cont vs cat comparisons 
# pc1 vs categorical comparisons 
exclude_strings_pca <- c('edss_cat_vs_pixel_cont_groups',
                         'edss_cat_vs_stride_cont_groups',
                         "t25fw_cat_vs_pixel_cont_groups",
                         't25fw_cat_vs_stride_cont_groups',
                         'edss_cat_vs_pc1_groups', 
                         't25fw_cat_vs_pc1_groups', 
                         'edss_cat_vs_either_vid_cont_groups', 
                         't25fw_cat_vs_either_vid_cont_groups'
                         )

table(plot_long_df_pca$cols_compared)

plot_long_df_pca_2 <- plot_long_df_pca %>%  
  filter(!grepl(paste(exclude_strings_pca, collapse="|"), cols_compared)) 

#%>% 
 # filter(!grepl('either', cols_compared)) # comparison of either video 

table(plot_long_df_pca_2$cols_compared)
view(plot_long_df_pca_2)
```

```{r}
# factor - level for plotting order 
print(table(plot_long_df_pca_2$video_metric))

plot_long_df_pca_2$video_metric = factor(plot_long_df_pca_2$video_metric, 
                                            levels = c('Pixel Proxy Continuous', 
                                                       'Stride Time Continuous',
                                                       'Either Continuous',
                                                       'PC1',
                                                       'Pixel Proxy Categorical', 
                                                       'Stride Time Categorical', 
                                                       'Either Categorical'
                                                       ))

print(table(plot_long_df_pca_2$clinical_outcome, plot_long_df_pca_2$video_metric)) # confirm

# factor clinical outcome variable - control order of facet_wrap
plot_long_df_pca_2$clinical_outcome <- factor(plot_long_df_pca_2$clinical_outcome, 
                                          levels = c('T25FW Continuous', 
                                                     'EDSS Ordinal', 
                                                     'T25FW Categorical',
                                                     'EDSS Categorical'))
```

### Plot with proportion 
```{r}
# add proportion column to put text on plot 
plot_long_df_pca_prop <- plot_long_df_pca_2 %>%
  group_by(clinical_outcome, video_metric, agreement_group) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(clinical_outcome, video_metric) %>%
  mutate(prop = n / sum(n),
         label = scales::percent(prop, accuracy = 1))

table(plot_long_df_pca_prop$video_metric)
```

```{r}
# proportion 
p <- ggplot(plot_long_df_pca_prop, aes(x = video_metric, y = prop, fill = agreement_group)) +
  geom_bar(stat = "identity", position = "fill") +
  geom_text(aes(label = ifelse(prop > 0.03, label, "")),  # Only label if > 5%
            position = position_fill(vjust = 0.5),
            size = 3.25) + 
  facet_grid( ~ clinical_outcome, scales = 'free_x') +
   scale_fill_brewer(name = "Worsened by:", palette = "Set2") +
  labs(y = "Proportion", x = "") +
  theme_minimal() + 
  theme(plot.title = element_text(size = 9, face = 'bold', hjust = 0.5),
          axis.title = element_text(size = 9),
          #axis.text = element_text(size = 9),
          axis.text.x = element_text(size = 9, face = 'italic', angle = 90),
          axis.text.y = element_blank(),
          axis.title.y = element_blank(),
          axis.ticks.y = element_blank(),
          legend.text = element_text(size = 9), 
          legend.title = element_text(size = 9),
          legend.position = 'bottom',
          legend.key.size = unit(0.5, "lines"))  
p
ggsave(file.path(long_output_dir, 'barcharts_proportion_3_wpc1.png'),
       bg = 'white', height = 5, width = 7)
```


